### 2017-11-19 doc/supplementary.tex
Slightly modified the notation used in Supplementary, so that the mini-batch can be easily incorporated.

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2016
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2016}

% \usepackage{nips_2016}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amsmath}	% for \begin{align}
\usepackage{graphicx}	% for \includegraphics{filename}
\usepackage{subcaption}	% for \begin{subfigure}[t]{0.5\textwidth}

\newcommand{\bb}[1]{\boldsymbol{#1}}

\title{Supplementary of Adversarial Zero-Shot Learning}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
	Shihui Li, Yu-Hsiang Lin, Kangyan Zhou
		%\thanks{Use footnote for providing further information about author (webpage, alternative address)---\emph{not} for acknowledging funding agencies.}
		\\
	Language Technologies Institute\\
	Carnegie Mellon University\\
	Pittsburgh, PA 15213 \\
	\texttt{\{shihuil,yuhsianl,kangyanz\}@andrew.cmu.edu} \\
	%% examples of more authors
	%% \And
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
	%% \And
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

%\begin{abstract}

%	This is a 4-page draft.

%\end{abstract}

% ----------------------------------------------------
% ----------------------------------------------------

\section{Adversarial autoencoder with the prior as the regularizer}

	Given an input image representation $\bb{x} \in R^{d_x}$, the encoder encodes it into $\bb{h} \in R^{d_h}$ by
	\begin{align}
		\bb{h} = \sigma(W^e \bb{x} + \bb{b}^e),
	\end{align}
	where $W^e \in R^{d_h \times d_x}$, $\bb{b}^e \in R^{d_h}$, and $\sigma(\cdot)$ is the sigmoid function,
	\begin{align}
		\sigma(s) = \frac{1}{1 + e^{-s}},
	\end{align}
	applied element-wisely. Note that the hidden representation space $R^{d_h}$ is also the space of the text representation; that is, $\bb{t} \in R^{d_h}$. The decoder computes
	\begin{align}
		\tilde{\bb{x}} = \sigma(W^d \bb{h} + \bb{b}^d),
	\end{align}
	where $W^d \in R^{d_x \times d_h}$ and $\bb{b}^d \in R^{d_x}$.

	The generator of the adversarial net is the encoder of the autoencoder; that is,
	\begin{align}
		G(\bb{x}) = \sigma(W^e \bb{x} + \bb{b}^e) = \bb{h}.
	\end{align}
	The positive samples are drawn from the Gaussian prior,
	\begin{align}
		\bb{z} \sim \mathcal{N}(0, \lambda^2).
	\end{align}
	where $\bb{z} \in R^{d_h}$.	We use a fully connected neural network with a single hidden layer as the discriminator. Given an input $\bb{z} \in R^{d_h}$, it computes
	\begin{align}
		\bb{z}^1 &= \sigma( W^1 \bb{z} + \bb{b}^1 ), \\
		%
		z^2 &= \sigma( (\bb{w}^2)^{\top} \bb{z}^1 + b^2 ),
	\end{align}
	where $\bb{z}^1 \in R^{d_1}$, $W^1 \in R^{d_1 \times d_h}$, $\bb{b}^1 \in R^{d_1}$, $\bb{w}^2 \in R^{d_1}$, and $b^2$ and $z^2$ are scalars. The discriminator returns
	\begin{align}
		D(\bb{z}) = z^2,
	\end{align}
	which is the estimation of the probability that $\bb{z}$ is drawn from the prior.

	For the autoencoder and the generator of the adversarial net, the loss function for a mini-batch of instances, $S$, is the empirical risk,
	\begin{align}
		f_g = \frac{1}{|S|} \sum_{i \in S} \left\{ -\bb{x}^{(i)} \odot \log \tilde{\bb{x}}^{(i)} + \log\left[ 1 - D( \bb{h}^{(i)} ) \right] + || \bb{h}^{(i)} - \bb{t}^{c_i} ||_2^2 \right\},
	\end{align}
	where $\odot$ denotes the element-wise product, the log function is applied element-wisely, and $c_i$ is the class of the instance $x^{(i)}$. Here we assume that there is only one text $\bb{t}^c$ for each class $c$.

	For the discriminator of the adversarial net, the loss function is
	\begin{align}
		f_d = -\frac{1}{|S|} \sum_{j = 1}^{|S|} \log D( \bb{z}^{(j)} ) - \frac{1}{|S|} \sum_{i \in S} \log\left[ 1 - D( \bb{h}^{(i)} ) \right],
	\end{align}
	where $\bb{z}^{(j)}$ are $|S|$ samples drawn from the prior distribution $\mathcal{N}(0, \lambda^2)$.

	The optimization problem for the autoencoder and the generator is
	\begin{align}
		\min_{W^e, \bb{b}^e, W^d, \bb{b}^d} \; f_g,
	\end{align}
	while the parameters of the discriminator ($W^1, \bb{b}^1, \bb{w}^2, b^2$) are held constant. The optimization problem for the discriminator is
	\begin{align}
		\min_{W^1, \bb{b}^1, \bb{w}^2, b^2} \; f_d,
	\end{align}
	while the parameters of the generator ($W^e, \bb{b}^e, W^d, \bb{b}^d$) are held constant.

% ----------------------------------------------------
% ----------------------------------------------------

%\bibliographystyle{plain}
%\bibliography{MachineLearning}

\end{document}